---
title: "Detection of Outlier SvH Scores - p-values"
author: "Robert Allaway"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    code_fold: hide
    toc: true
    toc_float: true
---

# Introduction 

In order to test the accuracy of the gold-standard data, we can leverage a wisdom-of-the-crowds approach where we compare the set of predicted values for a given measurement (e.g. all predicted values for Patient 001, Joint 001, erosion) to the gold standard measurement. We can calculate p-values for this population of predictions + gold standard value. If the p-value of the gold standard is considerably higher or lower than the rest of the population (i.e. very far from 0), it's possible that the gold standard measurement is incorrect. 

First, load packages and download View of all submissions. We'll consider only the final submission for each team to avoid weighting towards teams that submitted multiple similar predictions. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(tidyverse)
library(reticulate)
library(reactable)
use_condaenv("synapse-2") #conda environment with synapse >2.0 installed
synapse <- import('synapseclient')
syn <- synapse$Synapse()


```

```{r include=FALSE}
#login
syn$login('ra2dreamservice', '#####')

```

Then, calculate the p-values for each prediction and gold standard value, grouped by Patient_ID and measurement (i.e. p-values calculated within each group, not across every group).

```{r echo=TRUE, message=FALSE, warning=FALSE}

tab <- syn$tableQuery('select * from syn22236264')$filepath %>% 
      readr::read_csv() %>% 
  filter(status == "ACCEPTED") %>% 
  group_by(submitterid) %>% 
  top_n(1, createdOn) %>% 
  ungroup %>% 
  write_csv("submission_table.csv")

top_methods <- tab %>% top_frac(0.5, -sc1_weighted_sum_error)

submission <- lapply(top_methods$prediction_fileid, function(x){
  syn$get(x)$path %>% 
    readr::read_csv() %>% 
    tidyr::gather(measurement, score, -Patient_ID) %>% 
    mutate(prediction = {{x}})
}) %>% bind_rows() %>% write_csv("submissions.csv")

gold <- syn$get("syn22254942")$path %>% 
   readr::read_csv() %>% 
   tidyr::gather(measurement, score, -Patient_ID) %>% 
   rename(gold = score)

submission_pval<- left_join(submission, gold) %>% 
  group_by(Patient_ID, measurement) %>% 
  mutate(pvalue = t.test(score, mu = unique(gold))$p.value) %>% 
    mutate(mean_prediction = signif(mean(score),3)) %>% 
  select(Patient_ID, measurement, gold, pvalue, mean_prediction) %>% 
  distinct() %>% 
  ungroup() %>% 
  mutate(fdr = p.adjust(pvalue, method= "fdr")) %>% 
  mutate(pvalue = signif(pvalue, 3), fdr = signif(fdr, 3))

```

# SC1 

Plot the SC1 gold standard p-values. The p-values for these Overall_Tol values are not too high - I would surmise because these values are the sum of a large number of measurements, and thus are "buffered" from large changes caused by individual joint measurement errors. 

The table below shows the same data. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

p <-ggplot(submission_pval %>% 
         filter(measurement == "Overall_Tol")) +
  geom_point(aes(x = gold, y = fdr, text = Patient_ID, label = measurement)) +
  theme_bw() +
  ggtitle("SC1 Gold Standard p-values")
  
plotly::ggplotly(p)

p <-ggplot(submission_pval %>% 
         filter(measurement == "Overall_Tol")) +
  geom_point(aes(x = gold, y = mean_prediction, text = Patient_ID, label = measurement)) +
  theme_bw() +
  ggtitle("SC1 Gold Standard vs mean prediction")
  
plotly::ggplotly(p)
# submission_zscores %>% 
#          filter(prediction=="gold") %>% 
#          filter(measurement == "Overall_Tol") %>%
#         select(-prediction) %>% 
#   arrange(desc(zscore)) %>% 
#   reactable(sortable = T, filterable = F, bordered = T, compact = T,
#           style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"))

```

# SC2

Plot the SC2 gold standard p-values. This plot is not interactive because there are many data points. 
The table at the end of this document shows the same data for SC2 joint narrowing gold standard scores and is sortable.

```{r echo=TRUE, message=FALSE, warning=FALSE}

p <-ggplot(submission_pval %>% 
         filter(grepl('.+_J__.+', measurement)) %>% 
           filter(fdr < 0.05)) +
  geom_point(aes(x = gold, y = fdr, text = Patient_ID, label = measurement)) +
  theme_bw() +
  ggtitle("SC2 (narrowing) Gold Standard FDR p-values < 0.05 (x-axis jittered)")

p

p <-ggplot(submission_pval %>% 
         filter(grepl('.+_J__.+', measurement)) %>% 
           filter(fdr < 0.05)) +
  geom_jitter(aes(x = gold, y = mean_prediction, text = Patient_ID, label = measurement), width = 0.25) +
  theme_bw() +
  ggtitle("SC2 (narrowing) Gold Standard vs mean Prediction (x-axis jittered)")

p

submission_pval %>% 
  filter(grepl('.+_J__.+', measurement))  %>%
  arrange(fdr) %>% 
  reactable(sortable = T, filterable = F, bordered = T, compact = T,
          style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"))
```


# SC3

Plot the SC3 gold standard p-values. This plot is not interactive because there are many data points. 
The table at the end of this document shows the same data for SC3 joint narrowing gold standard scores and is sortable.

```{r echo=TRUE, message=FALSE, warning=FALSE}

p <-ggplot(submission_pval %>% 
         filter(grepl('.+_E__.+', measurement)) %>% 
           filter(fdr < 0.05)) +
  geom_point(aes(x = gold, y = fdr, text = Patient_ID, label = measurement)) +
  theme_bw() +
  ggtitle("SC3 (narrowing) Gold Standard FDR p-values < 0.05 (x-axis jittered)")

p

p <-ggplot(submission_pval %>% 
         filter(grepl('.+_E__.+', measurement)) %>% 
           filter(fdr < 0.05)) +
  geom_jitter(aes(x = gold, y = mean_prediction, text = Patient_ID, label = measurement), width = 0.25) +
  theme_bw() +
  ggtitle("SC3 (narrowing) Gold Standard vs mean Prediction (x-axis jittered)")

p

submission_pval %>% 
  filter(grepl('.+_E__.+', measurement))  %>%
  arrange(fdr) %>% 
  reactable(sortable = T, filterable = F, bordered = T, compact = T,
          style = list(fontFamily = "-apple-system, BlinkMacSystemFont, Segoe UI, Helvetica, Arial, sans-serif"))
```


```{r echo=TRUE, message=FALSE, warning=FALSE}
write_csv(submission_pval %>%  
          arrange(fdr),
  'gold_standard_p_values.csv')
```
