---
title: 'TEAR Validation - *adjusted* dataset - prediction scoring'
author: "Robert Allaway (Sage Bionetworks)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    df_print: paged
    code_fold: hide
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
---

## Introduction 

We have an independent dataset, TEAR, that also has hand and foot RA images. We tested the 15 final round docker containers on an independent TEAR cohort of 50 individuals to assess which methods are most stable in new datasets. For this experiment, we adjusted the dataset by normalizing the signal intensity for each image. This is because the original TEAR dataset had many images with poor contrast and high intensity, washing out many of the joint. Note: the gold standard scores for each image are from the original dataset, as we don't have any way of rescoring/adjusting those values. However, it's reasonable to assume that a trained observer, particularly with access to the original films, would be able to overcome the image issues in the original dataset more effectively than a model trained on a higher-quality dataset. The code used to adjust the files is available in this repository as "imadjust.py". 

## Setup

First, import packages for data manipulation and retrieve prediction data, gold standard data, and the template for use in the scoring code. Don't forget to set a seed!

```{r echo=TRUE, message=FALSE, warning=FALSE}
set.seed(98109)

library(tidyverse)
library(reticulate)
library(challengescoring)
library(ggplot2)
library(reactable)

# Synapse setup to use `reticulate`
use_condaenv("challenge", required = T)
synapseclient <- reticulate::import('synapseclient')
syn <- synapseclient$Synapse()
```

```{r include=FALSE}
#login
syn$login()

```

#### Functions 

Additionally, define functions to help with the scoring algorithms.

```{r echo=TRUE, message=FALSE, warning=FALSE}
# 
calculate_weight <- function(x){
  weight <- dplyr::if_else(x == 0, 1,
            dplyr::if_else(x == 1, 2,
            dplyr::if_else(x>=2 & x<=3, 2.143547,
            dplyr::if_else(x>=4 & x<=7, 3.863745,
            dplyr::if_else(x>=8 & x<=20, 8,
            dplyr::if_else(x>=21 & x<=55, 16,
            dplyr::if_else(x>=56 & x<=148, 32,
            dplyr::if_else(x>=148, 64, 0))))))))
  return(weight)
}

# 
sum_weighted_error <- function(gold, pred, weight){
  sum(weight*abs(gold - pred))
}

rmse <- function(gold, pred){
  sqrt(mean((gold - pred) ** 2))
}

score_log2 <- function(gold_path, prediction_path){
  
  pred <- readr::read_csv(prediction_path) %>% 
    dplyr::mutate_at(vars(-Patient_ID), ~ replace(., which(.<0), 0))
  gold <- readr::read_csv(gold_path) 
  
  ##score SC1 
  gold_sc1 <- gold %>% 
    dplyr::select(Patient_ID, Overall_Tol) %>% 
    dplyr::mutate(weight = calculate_weight(Overall_Tol))
  
  pred_sc1 <- pred %>% 
    dplyr::select(Patient_ID, Overall_Tol) %>% 
    dplyr::rename(Overall_Tol_pred = Overall_Tol)
  
  sc1_weighted_sum_error <- dplyr::full_join(pred_sc1, gold_sc1, by = "Patient_ID") %>% 
    dplyr::summarize(result = sum_weighted_error(log2(Overall_Tol+1), log2(Overall_Tol_pred+1), weight)/sum(weight)) %>% 
                       purrr::pluck("result")
                     
                     ##does sc2/3 prediction exist?
                     gold_joints <- gold %>% 
                       dplyr::mutate(weight = calculate_weight(Overall_Tol)) %>% 
                       dplyr::select(-Overall_Tol) %>% 
                       tidyr::gather('joint', 'score', -Patient_ID, -weight)
                     
                     pred_joints <- pred %>% 
                       dplyr::select(-Overall_Tol) %>% 
                       tidyr::gather('joint', 'pred_score', -Patient_ID)
                     
                     sc2_sc3 <- dplyr::full_join(pred_joints, gold_joints, by = c("Patient_ID","joint"))
                     
                     ##if participants set all joint values to 0, we will just return NA for sc2/3
                     if(!all(pred_joints$pred_score == 0)){
                       
                       #score sc2
                       sc2_total_weighted_sum_error <- sc2_sc3 %>% 
                         dplyr::filter(grepl('Overall_erosion', joint)) %>%
                         dplyr::summarize(result = sum_weighted_error(log2(score+1), log2(pred_score+1), weight)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc2_joint_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl(".+_E__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc2_hand_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl("[RL]H.+_E__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc2_foot_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl("[RL]F.+_E__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       ##score SC3
                       sc3_total_weighted_sum_error <- sc2_sc3 %>% 
                         dplyr::filter(grepl('Overall_narrowing', joint)) %>%
                         dplyr::summarize(result = sum_weighted_error(log2(score+1), log2(pred_score+1), weight)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc3_joint_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl(".+_J__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc3_hand_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl("[RL]H.+_J__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       sc3_foot_weighted_sum_rmse <- sc2_sc3 %>% 
                         dplyr::filter(grepl("[RL]F.+_J__.+", joint)) %>% 
                         dplyr::group_by(Patient_ID, weight) %>% 
                         dplyr::summarize(patient_rmse = rmse(log2(score+1), log2(pred_score+1))) %>% 
                         dplyr::ungroup() %>% 
                         dplyr::summarize(result= sum(weight*patient_rmse)/sum(weight)) %>% 
                         purrr::pluck("result")
                       
                       
                     }else{
                       sc2_total_weighted_sum_error <- NA
                       sc2_joint_weighted_sum_rmse <- NA
                       sc2_hand_weighted_sum_rmse <- NA 
                       sc2_foot_weighted_sum_rmse <- NA
                       sc3_total_weighted_sum_error <- NA
                       sc3_joint_weighted_sum_rmse <- NA
                       sc3_hand_weighted_sum_rmse <- NA
                       sc3_foot_weighted_sum_rmse <- NA 
                     }
                     
                     score <- c(
                       "sc1_weighted_sum_error" = sc1_weighted_sum_error,
                       "sc2_total_weighted_sum_error" = sc2_total_weighted_sum_error,
                       "sc2_joint_weighted_sum_rmse" = sc2_joint_weighted_sum_rmse,
                       "sc2_hand_weighted_sum_rmse" = sc2_hand_weighted_sum_rmse, 
                       "sc2_foot_weighted_sum_rmse" = sc2_foot_weighted_sum_rmse,
                       "sc3_total_weighted_sum_error" = sc3_total_weighted_sum_error,
                       "sc3_joint_weighted_sum_rmse" = sc3_joint_weighted_sum_rmse,
                       "sc3_hand_weighted_sum_rmse" = sc3_hand_weighted_sum_rmse, 
                       "sc3_foot_weighted_sum_rmse" = sc3_foot_weighted_sum_rmse
                     )
                     
                     return(score)
                     
}


```

#### Goldstandard

The assessment of subchallenge 1 utilizies known SvH scores and subchallenges 2 & 3 known individual joint narrowing and erosion scores. After adding a weight for each joint score, split the goldstandard accordingly. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
gold_1 <- syn$get("syn25324263")$path

gold_2 <- syn$get("syn25324262")$path
```


```{r echo=TRUE, message=FALSE, warning=FALSE}

query <- syn$tableQuery(
  "SELECT id, prediction_fileid, orginial_submitterid as original_submitterid, createdOn,
    sc1_weighted_sum_error AS sc1, 
    sc2_joint_weighted_sum_rmse AS sc2, 
    sc3_joint_weighted_sum_rmse AS sc3
  FROM syn25710243 WHERE status = 'ACCEPTED' and 
    orginial_submitterid <> 3408914 AND
    orginial_submitterid <> 3399952 AND
  dockerdigest not in ('sha256:6cb365830745c5d368bcd128a3a1b750ace801da281356a5a623237408c10a3d')")$asDataFrame() 

# For easier identification, replace each team/participant's submitterid with
# their team name/username.
query$original_submitterid <- as.character(query$original_submitterid)

team_names <- sapply(query$original_submitterid, function(sub) {
  name <- tryCatch({
    syn$getUserProfile(sub)$userName
  }, error = function(err) {
    syn$getTeam(sub)$name
  })
  return(name)
})

query$original_submitterid <- team_names

pred_filenames <- lapply(query$prediction_fileid, function(id) {
  syn$get(id)$path
})

names(pred_filenames) <- team_names

# Before bootstrapping, rearrange the teams by rank, where left = better 
# performance and right = not-as-great performance.

### SC1
test1_scores <- sapply(pred_filenames, score_log2, gold_path = gold_1) %>% 
  t() %>% 
  as_tibble(rownames = "team") %>% 
  tidyr::pivot_longer(names_to = "category", values_to = "score_1", cols = -c("team"))

test2_scores <- sapply(pred_filenames, score_log2, gold_path = gold_2) %>% 
  t() %>% 
  as_tibble(rownames = "team") %>%
  tidyr::pivot_longer(names_to = "category", values_to = "score_2", cols = -c("team"))

all_scores <- full_join(test1_scores, test2_scores) %>% 
  filter(category %in% c("sc1_weighted_sum_error", "sc2_joint_weighted_sum_rmse", "sc3_joint_weighted_sum_rmse"))

ggplot(data = all_scores) +
  geom_point(aes(x = score_1, y = score_2)) +
    ggrepel::geom_text_repel(aes(x= score_1, y = score_2, label = team), size = 2, force = 200, box.padding = 0.1, min.segment.length = 0, direction = "both", segment.colour = "darkgray") +
  facet_wrap(category ~ ., scales = 'free') + 
  labs(x = "Gold standard one", y = "Gold standard two") +
  theme_bw()

ggsave("tear_adjusted_dual_gold_standards.pdf", width = 10, height = 4)

```


```{r echo=TRUE, message=FALSE, warning=FALSE}

mean_validation <- all_scores %>% 
  filter(category %in% c("sc1_weighted_sum_error", "sc2_joint_weighted_sum_rmse", "sc3_joint_weighted_sum_rmse")) %>% 
  mutate(mean_validation_score = ((score_1+score_2)/2))

##get final round submissions
query <- syn$tableQuery(
  "SELECT id, prediction_fileid, submitterid, createdOn,
    sc1_weighted_sum_error, 
    sc2_joint_weighted_sum_rmse, 
    sc3_joint_weighted_sum_rmse
  FROM syn22236264 WHERE status = 'ACCEPTED' AND 
    submitterid <> 3408914 AND
    submitterid <> 3399952 AND
    dockerdigest not in ('sha256:6cb365830745c5d368bcd128a3a1b750ace801da281356a5a623237408c10a3d')")$asDataFrame() %>%
  group_by(submitterid) %>%
  slice(which.max(createdOn)) %>%
  select(-createdOn)  
  
query$submitterid <- as.character(query$submitterid)

team_names <- sapply(query$submitterid, function(sub) {
  name <- tryCatch({
    syn$getUserProfile(sub)$userName
  }, error = function(err) {
    syn$getTeam(sub)$name
  })
  return(name)
})

query$team <- team_names

original_scores <- query %>% 
  ungroup() %>% 
  select(team, sc1_weighted_sum_error, sc2_joint_weighted_sum_rmse, sc3_joint_weighted_sum_rmse) %>% 
  tidyr::pivot_longer(names_to = "category", values_to = "score_orig", cols = -c("team"))

valid_vs_original <- mean_validation %>% 
  select(team, category, mean_validation_score) %>% 
  full_join(original_scores) %>% 
  filter(team != 'vladyorsh') %>%  ## don't have validation predictions for this submitter
  mutate(sc = case_when(category == "sc1_weighted_sum_error" ~ "Subchallenge 1",
                   category == "sc2_joint_weighted_sum_rmse" ~ "Subchallenge 2",
                   category == "sc3_joint_weighted_sum_rmse" ~ "Subchallenge 3"))




plts <- lapply(unique(valid_vs_original$sc), function(x){
  if(x == "Subchallenge 1"){
    lims <- c(0, 4.75)
  }else{
    lims <- c(0,1)
  }
  
ggplot(data = valid_vs_original %>% filter(sc %in% x)) +
  geom_point(aes(x = score_orig, y = mean_validation_score), size = 0.75, alpha = 0.5) +
  labs(x = "Final round score", y = "Post-challenge validation score (mean)") +
  xlim(lims) +
  ylim(lims) +
  theme_bw() + 
  geom_hline(data = valid_vs_original %>% 
                   filter(sc %in% x) %>% 
                   filter(team == "RA2 Baseline Model") %>% distinct(), 
             aes(yintercept = mean_validation_score),
             color = '#DD0426',
             alpha = 0.25,
             linetype = "dashed",
             size = 1) + ##shows whether better than baseline 
  geom_vline(data = valid_vs_original %>% 
                   filter(sc %in% x) %>% 
                   filter(team == "RA2 Baseline Model") %>% distinct(), 
             aes(xintercept = score_orig),
             color = '#FFA62B',
             alpha = 0.5,
             linetype = "dashed",
             size = 1) + ##shows whether better than baseline 
  geom_abline(slope = 1,              
             color = '#009FB7',
             alpha = 0.5,
             linetype = "dashed",
             size = 1) + ##shows whether better across final vs validation rnd
  ggrepel::geom_text_repel(aes(x= score_orig, y = mean_validation_score, label = team), size = 2, force = 10, point.padding = 0.05, box.padding = 0.5, segment.colour = "darkgray", max.overlaps = 100, min.segment.length = 0.001, direction = "both") 
})

cowplot::plot_grid(plotlist = plts, rows = 1)
ggsave("tear_adjusted_vs_final_round.pdf", width = 12, height = 4)

valid_vs_original_gathered <-  valid_vs_original %>% 
  tidyr::pivot_longer(names_to = "orig_vs_valid", values_to = "score", cols = -c("team","category","sc")) %>% 
  mutate(label = case_when(orig_vs_valid == "mean_validation_score" ~ "Validation \n score (mean)",
                   orig_vs_valid == "score_orig" ~ "Final round \n score")) %>% 
  group_by(label, sc) %>%
  mutate(ranks = rank(score))

ord <- c("Final round \n score", "Validation \n score (mean)")

     
ggplot(data = valid_vs_original_gathered,
       aes(x = fct_relevel(label, ord), y = ranks, color = team)) +
  geom_point() +
  geom_text(data = filter(valid_vs_original_gathered, orig_vs_valid == 'mean_validation_score'),
            aes(x = fct_relevel(label, ord), y = ranks, label = team), hjust = "left", nudge_x = 0.05, size = 2.4) +
  geom_line(aes(group = team))+
  facet_wrap(. ~ sc, nrow = 1, scales = 'free') + 
  labs(x = "Round", y = "Rank") +
  shades::hue(scale_color_viridis_d("Team", option = "C", direction = -1), shades::scalefac(0.80)) +
  theme_bw() +
  scale_y_continuous(labels = scales::label_ordinal(), breaks = scales::breaks_width(1)) +
  scale_x_discrete(expand = expansion(mult = c(.2, 1.6)))

ggsave("tear_adjusted_vs_final_round_lineplot.pdf", width = 12, height = 4)


```

